---
title: "DeepLearning-6"
author: "Tinniam V Ganesh"
date: "April 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
```

#Introduction
In this 6th instalment of 'Deep Learning from first principles in Python, R and Octave-Part6', I look at a couple of different initialization techniques used in Deep Learning, L2 regularization and the 'dropout' method. Specifically, I implement "He initialization" & "Xavier Initialization". My earlier posts in this series of Deep Learning included

1. [Part 1](https://gigadom.wordpress.com/2018/01/04/deep-learning-from-basic-principles-in-python-r-and-octave-part-1/) - In the 1st part I implemented logistic regression as a simple 2 layer Neural Network
2. [Part 2](https://gigadom.wordpress.com/2018/01/11/deep-learning-from-first-principles-in-python-r-and-octave-part-2/) - In part 2, implemented the most basic of Neural Networks, with just 1 hidden layer, and any number of activation units in that hidden layer. The implementation was in vectorized Python, R and Octave
3. [Part 3](https://gigadom.wordpress.com/2018/01/30/deep-learning-from-first-principles-in-python-r-and-octave-part-3/) -In part 3, I derive the equations and also implement a L-Layer Deep Learning network with either the relu, tanh or sigmoid activation function in Python, R and Octave. The output activation unit was a sigmoid function for logistic classification
4. [Part 4](https://gigadom.wordpress.com/2018/02/26/deep-learning-from-first-principles-in-python-r-and-octave-part-4/) - This part looks at multi-class classification, and I derive the Jacobian of a Softmax function and implement a simple problem to perform multi-class classification.
5. [Part 5](https://gigadom.wordpress.com/2018/03/23/deep-learning-from-first-principles-in-python-r-and-octave-part-5/) - In the 5th part, I extend the L-Layer Deep Learning network implemented in Part 3, to include the Softmax classification. I also use this L-layer implementation to classify MNIST handwritten digits with Python, R and Octave.

The code in Python, R and Octave are identical, and just take into some of the minor idiosyncrasies of the individual language

##1. Initialzation techniques
The usual initialization technique is to generate Gaussian or uniform random numbers and multiply it by a small value like 0.01. Two techniques which are used to speed up convergence is the [He](https://www.cv-foundation.org/openaccess/content_iccv_2015/papers/He_Delving_Deep_into_ICCV_2015_paper.pdf)" initialization or [Xavier](http://proceedings.mlr.press/v9/glorot10a/glorot10a.pdf). 

**Random Initialization**
This technique just initializes the weights to small random values based on Gaussian or uniform distribution


```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())
#Load the data
train_X, train_Y, test_X, test_Y = load_dataset()
# Set the layers dimensions
layersDimensions = [2,7,1]

# Train a deep learning network with random initialization
parameters = L_Layer_DeepModel(train_X, train_Y, layersDimensions, hiddenActivationFunc='relu', outputActivationFunc="sigmoid",learningRate = 0.6, num_iterations = 9000, initType="default", print_cost = True,figure="fig1.png")

# Clear the plot
plt.clf()
plt.close()

# Plot the decision boundary
plot_decision_boundary(lambda x: predict(parameters, x.T), train_X, train_Y,str(0.6),figure1="fig2.png")
```


**He Initialization**
He initialization multiply the random weights by \sqrt{\frac{2}{dimension\ of\ previous\ layer}}

```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())

#Load the data
train_X, train_Y, test_X, test_Y = load_dataset()
# Set the layers dimensions
layersDimensions = [2,7,1]

# Train a deep learning network with He  initialization
parameters = L_Layer_DeepModel(train_X, train_Y, layersDimensions, hiddenActivationFunc='relu', outputActivationFunc="sigmoid", learningRate =0.6,    num_iterations = 10000,initType="He",print_cost = True,                           figure="fig3.png")

plt.clf()
plt.close()
# Plot the decision boundary
plot_decision_boundary(lambda x: predict(parameters, x.T), train_X, train_Y,str(0.6),figure1="fig4.png")
```


**Xavier Initialization**
Xavier initialization multiply the random weights by \sqrt{\frac{1}{dimension\ of\ previous\ layer}

```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())

#Load the data
train_X, train_Y, test_X, test_Y = load_dataset()
# Set the layers dimensions
layersDimensions = [2,7,1]
 
parameters = L_Layer_DeepModel(train_X, train_Y, layersDimensions, hiddenActivationFunc='relu', outputActivationFunc="sigmoid",
                            learningRate = 0.6,num_iterations = 10000, initType="Xavier",print_cost = True,
                            figure="fig5.png")

# Plot the decision boundary
plot_decision_boundary(lambda x: predict(parameters, x.T), train_X, train_Y,str(0.6),figure1="fig6.png")
```


## Default initialization - R
```{r fig1, cache=TRUE}
source("DLfunctions61.R")
z <- as.matrix(read.csv("circles.csv",header=FALSE)) 

x <- z[,1:2]
y <- z[,3]
X <- t(x)
Y <- t(y)
layersDimensions = c(2,11,1)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="sigmoid",
                            learningRate = 0.5,
                            numIterations = 9000, 
                            initType="default",
                            print_cost = True)
#Plot the cost vs iterations
iterations <- seq(0,9000,1000)
costs=retvals$costs
df=data.frame(iterations,costs)
ggplot(df,aes(x=iterations,y=costs)) + geom_point() + geom_line(color="blue") +
 ggtitle("Costs vs iterations") + xlab("No of iterations") + ylab("Cost")

# Plot the decision boundary
plotDecisionBoundary(z,retvals,hiddenActivationFunc="relu",lr=0.5)

```



## He initialization
```{r fig 2,cache=TRUE}
source("DLfunctions61.R")
z <- as.matrix(read.csv("circles.csv",header=FALSE)) 

x <- z[,1:2]
y <- z[,3]
X <- t(x)
Y <- t(y)
layersDimensions = c(2,11,1)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="sigmoid",
                            learningRate = 0.5,
                            numIterations = 9000, 
                            initType="He",
                            print_cost = True)

#Plot the cost vs iterations
iterations <- seq(0,9000,1000)
costs=retvals$costs
df=data.frame(iterations,costs)
ggplot(df,aes(x=iterations,y=costs)) + geom_point() + geom_line(color="blue") +
    ggtitle("Costs vs iterations") + xlab("No of iterations") + ylab("Cost")

# Plot the decision boundary
plotDecisionBoundary(z,retvals,hiddenActivationFunc="relu",0.5,lr=0.5)

```

## Xavier initialization
```{r fig3, cache=TRUE}
## Xav initialization 
layersDimensions = c(2,11,1)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="sigmoid",
                            learningRate = 0.5,
                            numIterations = 9000, 
                            initType="Xav",
                            print_cost = True)

#Plot the cost vs iterations
iterations <- seq(0,9000,1000)
costs=retvals$costs
df=data.frame(iterations,costs)
ggplot(df,aes(x=iterations,y=costs)) + geom_point() + geom_line(color="blue") +
    ggtitle("Costs vs iterations") + xlab("No of iterations") + ylab("Cost")

# Plot the decision boundary
plotDecisionBoundary(z,retvals,hiddenActivationFunc="relu",0.5)
```


##Regularization - Python

```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())

#Load the data
train_X, train_Y, test_X, test_Y = load_dataset()
# Set the layers dimensions
layersDimensions = [2,7,1]

# Train a deep learning network
parameters = L_Layer_DeepModel(train_X, train_Y, layersDimensions, hiddenActivationFunc='relu',  
                               outputActivationFunc="sigmoid",learningRate = 0.6, lambd=0.1, num_iterations = 9000, 
                               initType="default", print_cost = True,figure="fig7.png")

# Clear the plot
plt.clf()
plt.close()

# Plot the decision boundary
plot_decision_boundary(lambda x: predict(parameters, x.T), train_X, train_Y,str(0.6),figure1="fig8.png")


plt.clf()
plt.close()
plot_decision_boundary(lambda x: predict(parameters, x.T,keep_prob=0.9), train_X, train_Y,str(2.2),"fig8.png",)
```


## Spiral data Regularization 2 - Python
```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())
N = 100 # number of points per class
D = 2 # dimensionality
K = 3 # number of classes
X = np.zeros((N*K,D)) # data matrix (each row = single example)
y = np.zeros(N*K, dtype='uint8') # class labels
for j in range(K):
  ix = range(N*j,N*(j+1))
  r = np.linspace(0.0,1,N) # radius
  t = np.linspace(j*4,(j+1)*4,N) + np.random.randn(N)*0.2 # theta
  X[ix] = np.c_[r*np.sin(t), r*np.cos(t)]
  y[ix] = j


# Plot the data
plt.scatter(X[:, 0], X[:, 1], c=y, s=40, cmap=plt.cm.Spectral)
plt.clf()
plt.close()  
layersDimensions = [2,100,3]
y1=y.reshape(-1,1).T
parameters = L_Layer_DeepModel(X.T, y1, layersDimensions, hiddenActivationFunc='relu', outputActivationFunc="softmax",
                           learningRate = 0.6,lambd=0.2, num_iterations = 5000, print_cost = True,figure="fig9.png")

plt.clf()
plt.close()  
W1=parameters['W1']
b1=parameters['b1']
W2=parameters['W2']
b2=parameters['b2']
plot_decision_boundary1(X, y1,W1,b1,W2,b2,figure2="fig10.png")
```


```{r fig4, cache=TRUE}
source("DLfunctions61.R")
df=read.csv("circles.csv",header=FALSE)

z <- as.matrix(read.csv("circles.csv",header=FALSE)) 

x <- z[,1:2]
y <- z[,3]
X <- t(x)
Y <- t(y)
layersDimensions = c(2,11,1)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="sigmoid",
                            learningRate = 0.5,
                            lambd=0.1,
                            numIterations = 9000, 
                            initType="default",
                            print_cost = True)

#Plot the cost vs iterations
iterations <- seq(0,9000,1000)
costs=retvals$costs
df=data.frame(iterations,costs)
ggplot(df,aes(x=iterations,y=costs)) + geom_point() + geom_line(color="blue") +
    ggtitle("Costs vs iterations") + xlab("No of iterations") + ylab("Cost")


# Plot the decision boundary
plotDecisionBoundary(z,retvals,hiddenActivationFunc="relu",0.5)
```


## Regularization 2
```{r fig5, cache=TRUE}
# Read the spiral dataset
source("DLfunctions61.R")
Z <- as.matrix(read.csv("spiral.csv",header=FALSE)) 

# Setup the data
X <- Z[,1:2]
y <- Z[,3]
X <- t(X)
Y <- t(y)

layersDimensions = c(2, 15, 6, 3) #lr 2.6 20K, 0.15
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="softmax",
                            learningRate = 5.1,
                            lambd=0.1,
                            numIterations = 9000, 
                            print_cost = True)



parameters<-retvals$parameters
plotDecisionBoundary1(Z,parameters)
```

## Dropout 1 - Python
```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())
#Load the data
train_X, train_Y, test_X, test_Y = load_dataset()
# Set the layers dimensions
layersDimensions = [2,7,1]

# Train a deep learning network
parameters = L_Layer_DeepModel(train_X, train_Y, layersDimensions, hiddenActivationFunc='relu',  
                               outputActivationFunc="sigmoid",learningRate = 0.6, keep_prob=0.7, num_iterations = 9000, 
                               initType="default", print_cost = True,figure="fig11.png")

# Clear the plot
plt.clf()
plt.close()

# Plot the decision boundary
plot_decision_boundary(lambda x: predict(parameters, x.T,keep_prob=0.7), train_X, train_Y,str(0.6),figure1="fig12.png")

```


### Dropout 2 - Python
```{python}
import numpy as np
import matplotlib
import matplotlib.pyplot as plt
import sklearn.linear_model
import pandas as pd
import sklearn
import sklearn.datasets
exec(open("DLfunctions61.py").read())
# Create an input data set - Taken from CS231n Convolutional Neural networks,
# http://cs231n.github.io/neural-networks-case-study/
               

N = 100 # number of points per class
D = 2 # dimensionality
K = 3 # number of classes
X = np.zeros((N*K,D)) # data matrix (each row = single example)
y = np.zeros(N*K, dtype='uint8') # class labels
for j in range(K):
  ix = range(N*j,N*(j+1))
  r = np.linspace(0.0,1,N) # radius
  t = np.linspace(j*4,(j+1)*4,N) + np.random.randn(N)*0.2 # theta
  X[ix] = np.c_[r*np.sin(t), r*np.cos(t)]
  y[ix] = j


# Plot the data
plt.scatter(X[:, 0], X[:, 1], c=y, s=40, cmap=plt.cm.Spectral)
plt.clf()
plt.close()  
layersDimensions = [2,100,3]
y1=y.reshape(-1,1).T
parameters = L_Layer_DeepModel(X.T, y1, layersDimensions, hiddenActivationFunc='relu', outputActivationFunc="softmax",
                           learningRate = 0.6,keep_prob=0.5, num_iterations = 5000, print_cost = True,figure="fig13.png")

plt.clf()
plt.close()  
W1=parameters['W1']
b1=parameters['b1']
W2=parameters['W2']
b2=parameters['b2']
plot_decision_boundary1(X, y1,W1,b1,W2,b2,figure2="fig14.png")
```


## Dropout - R
```{r fig7, cache=TRUE}
source("DLfunctions61.R")
df=read.csv("circles.csv",header=FALSE)

z <- as.matrix(read.csv("circles.csv",header=FALSE)) 

x <- z[,1:2]
y <- z[,3]
X <- t(x)
Y <- t(y)
layersDimensions = c(2,11,1)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="sigmoid",
                            learningRate = 0.5,
                            keep_prob=0.8,
                            numIterations = 9000, 
                            initType="default",
                            print_cost = True)

# Plot the decision boundary
plotDecisionBoundary(z,retvals,keep_prob=0.6, hiddenActivationFunc="relu",0.5)
```



```{r fig8, cache=TRUE}
# Read the spiral dataset
source("DLfunctions61.R")
Z <- as.matrix(read.csv("spiral.csv",header=FALSE)) 

# Setup the data
X <- Z[,1:2]
y <- Z[,3]
X <- t(X)
Y <- t(y)

source("DLfunctions61.R")
layersDimensions = c(2, 15, 6, 3)
retvals = L_Layer_DeepModel(X, Y, layersDimensions,
                            hiddenActivationFunc='relu',
                            outputActivationFunc="softmax",
                            learningRate = 5.1,
                            keep_prob=0.8,
                            numIterations = 9000, 
                            print_cost = True)

parameters<-retvals$parameters
plotDecisionBoundary1(Z,parameters)

```

